%\documentclass[draftdoc]{JSBMLdoc}
\documentclass{JSBMLdoc}

\newcommand{\jsbmlversion}{1.2}
\newcommand{\docdate}{\today}

\makeindex

\input{../common/tex/macros.tex}
\input{../common/tex/hyphenation.tex}

\hypersetup{
  bookmarksopenlevel={1},
  bookmarksnumbered={true},
  breaklinks={true},
  pdfpagemode={UseOutlines},
  pdftitle={User guide for JSBML},
  pdfauthor={Andreas Dr\"ager} {Nicolas Rodriguez} {Alex Thomas} {Marine Dumousseau}
 {Alexander D\"orr} {Clemens Wrzodek} {Finja B\"uchel}
{Florian Mittag} {Nicolas Le Nov{\`e}re} {Andreas Zell} {Michael Hucka},
  pdfsubject={Software guide},
  pdfkeywords={JSBML} {libSBML} {Java} {SBML} {API} {LaTeX} {documentation}
{manual} {guide} {code examples},
  pdfview={FitV},
  pdffitwindow={true},
  pdfstartview={FitV},
  pdfnewwindow={false},
  pdfdisplaydoctitle={true},
  pdfhighlight={/P},
  plainpages={false},
  unicode={true}
}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{ %
  language=Java,                  % the language of the code
%  basicstyle=\footnotesize,       % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  stepnumber=1,                   % the step between two line-numbers. If it's 1, each line 
                                  % will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,                   % adds a frame around the code
  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=4,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                 % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},          % keyword style
  commentstyle=\color{dkgreen},       % comment style
  stringstyle=\color{mauve},         % string literal style
  escapeinside={|}{|},            % if you want to add a comment within your code
  morekeywords={*,...}               % if you want to add more keywords to the set
}

% -----------------------------------------------------------------------------
\begin{document}

\title{\textls[20]{JSBML User Guide}}

\version{\jsbmlversion\\[0.5em]{\normalsize\emph{Document build date: \docdate}}}

\newcommand{\where}[1]{\,\textsuperscript{#1}}
\newcommand{\divider}[1]{\multicolumn{3}{c}{\emph{#1}:}}

\author{%
  \setlength{\tabcolsep}{20pt}%
  \begin{tabular}{@{}ccc@{}}
    \divider{Authors}\\[0.75em]
    Andreas Dr\"ager\where{a,b}   & Nicolas Rodriguez\where{c,d} & Alex Thomas\where{a}\\
    Marine Dumousseau\where{d}    & Alexander D\"orr\where{b}    & Clemens Wrzodek\where{b}\\
    Finja B\"uchel\where{b}       & Florian Mittag\where{b}      & Michael Hucka\where{e}\\[2em]
    \divider{Principal investigators}\\[0.75em]
    Bernhard \O. Palsson\where{a} & Andreas Zell\where{b}        & Nicolas Le Nov\`ere\where{c,d}\\
                                  & Michael Hucka\where{e}\\[2em]
    \divider{Institutional affiliations}\\[0.5em]
  \end{tabular}
  \\
  \begin{normalsize}
  \where{a\,}Systems Biology Research Group, University of California, San Diego, La Jolla, CA, USA
  \\[0.25em]
  \where{b\,}Center for Bioinformatics Tuebingen, University of Tuebingen, T\"ubingen, Germany
  \\[0.25em]
  \where{c\,}The Babraham Institute, Babraham Campus, Cambridge, UK
  \\[0.25em]
  \where{d\,}European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridge, UK
  \\[0.25em]
  \where{e\,}Computing and Mathematical Sciences, California Institute of Technology, Pasadena, CA, USA
  \end{normalsize}
}

\frontNotice{SBML (the Systems Biology Markup Language) is an XML-based
  model representation format for storing and exchanging computational
  descriptions of biological processes.  To read, write, manipulate, and
  perform higher-level operations on SBML files and data streams, software
  applications need to map SBML entities to suitable software objects.
  JSBML provides a pure Java library for this purpose.  It supports all
  Levels and Versions of SBML and provides many powerful features,
  including facilities to help migrate from the use of libSBML (a popular
  library for SBML that is not written in Java).
  \\ \\
  This document provides an introduction to JSBML and its use.  It is aimed
  at both developers writing new Java-based applications as well as those
  who want to adapt libSBML-based applications to using JSBML.  This user
  guide is a companion to the JSBML API documentation.
  \\ \\
  \centerline{The JSBML home page is \url{\jsbmlHomePageURL}.}\\
  \centerline{The JSBML discussion group is \url{\jsbmlGroupURL}.}  }


\maketitlepage
\maketableofcontents
\clearpage


\chapter{Getting started with JSBML}
\label{chp:getting-started}

JSBML is a Java\TTra library that will help you to read, write and
manipulate SBML (Systems Biology Markup Language)
files~\cite{Draeger2011a, Draeger2011b}. This chapter provides
information for quickly getting started with using JSBML.

Before you can use JSBML, you will need to obtain a copy of the library.
\sec{sec:obtaining-jsbml} below describes different ways of doing
this, and explains which additional libraries you may need. JSBML also
requires the use of a Java Runtime Environment (JRE) version~1.6 or
later~\cite{JavaDownloadURL}. \index{Java Runtime Environment (JRE)} In the
rest of this document, we assume that you have already installed a suitable
JRE or Java Development Kit (JDK), and know how to configure the Java class
path on your system. \index{Java Development Kit (JDK)}

It is also essential to \emph{understand SBML} in order to be able to use
it (and JSBML) properly. If you are not already familiar with SBML, a good
starting point for learning about it is the latest SBML
specification~\cite{SBMLspecs}. You can find answers to many questions in
the SBML FAQ~\cite{SBMLFAQ} and optionally by asking on one of the SBML
discussion lists~\cite{SBMLforums}.

\input{JSBML_getting_started}
\input{JSBML_hello_world_examples}
\input{JSBML_more_examples}


\chapter{Differences between JSBML and libSBML}
\label{chp:jsbml-libsbml-diffs}

Prior to the availability of JSBML, the most widely-used API library for
SBML offering a Java interface has been libSBML~\cite{Bornstein2008}. As a
result, many Java application developers working with SBML are already
accustomed to the classes, methods and general approach provided by
libSBML. This chapter discusses the main differences between these two
libraries, and is aimed at current libSBML users who want to transition to
using JSBML. We also provide some programming examples and hints for how
to use and work with JSBML. In addition, we provide an overview of the type hierarchy 
and API of JSBML.

\input{JSBML_libSBML_comparison}
\input{JSBML_libSBML_class_differences}
\input{JSBML_libSBML_API_differences}


\chapter{Additional features provided by JSBML}
\label{chp:additional-jsbml-features}

The previous chapter covered many features of the JSBML API and how they
compare to those provided by libSBML's API.  In addition to the features
described in that chapter, JSBML also provides a number of capabilities that
are not found in libSBML.  This chapter briefly introduces the most important
additional capabilities.

\input{JSBML_change_listeners}
\input{JSBML_overdetermination_validator}
\input{JSBML_find_methods}
\input{JSBML_utility_classes}
\input{JSBML_logging_facilities}
\input{JSBML_modules}
\input{JSBML_offline_validation}


\chapter{Implementing extensions in JSBML}
\label{chp:howToExtension}

In this chapter, we describe how to get started with writing an extension
for JSBML to support an \SBMLthree package.  We use a concrete (though
artificial) example to illustrate various points.  This example extension
is named, very cleverly, \emph{Example}, and while it does not actually do
anything significant, we hope it will help make the explanations more
understandable. This chapter applies to JSBML version 1.0 only;
the 0.8 branch of JSBML does not support extension packages.

\input{JSBML_extensions_organization}
\input{JSBML_extensions_creating_objects}
\input{JSBML_extensions_reader_writer}
\input{JSBML_extensions_implementation_checklist}
\input{JSBML_extensions_eclipse_templates}
\input{JSBML_extensions_overview}


\chapter{Acknowledgments}
\label{chp:acknowledgements}

\input{JSBML_Acknowledgments}

% -----------------------------------------------------------------------------
% Appendix
% -----------------------------------------------------------------------------

\appendix

\chapter{Frequently Asked Questions (FAQ)}
\label{chp:faq}

\input{JSBML_FAQ}

\chapter{Open tasks in JSBML development}
\label{chp:open-tasks}

\input{JSBML_open_tasks}


% -----------------------------------------------------------------------------
% References
% -----------------------------------------------------------------------------

\clearpage

\thispagestyle{plain}
\pagestyle{plain}
\bibliography{../common/tex/literature}

% -----------------------------------------------------------------------------
% Index
% -----------------------------------------------------------------------------

\setindexprenote{\vspace*{0.1ex}}
\printindex


\end{document}


%%% Local Variables: 
%%% fill-column: 75
%%% End: 
