\documentclass[draftdoc]{JSBMLdoc}
\usepackage{dirtree}
\usepackage{makeidx}
\usepackage[font=small,totoc]{idxlayout}

\makeindex

\input{../common/tex/macros.tex}
\input{../common/tex/hyphenation.tex}

\hypersetup{
  bookmarksopenlevel={1},
  bookmarksnumbered={true},
  breaklinks={true},
  pdfpagemode={UseOutlines},
  pdftitle={User guide for JSBML},
  pdfauthor={Andreas Dr\"ager} {Nicolas Rodriguez} {Marine Dumousseau}
{Alexander D\"orr} {Clemens Wrzodek} {Finja B\"uchel} {Florian Mittag}
{Nicolas Le Nov{\`e}re} {Andreas Zell} {Michael Hucka},
  pdfsubject={Software guide},
  pdfkeywords={JSBML} {libSBML} {Java} {SBML} {API} {LaTeX} {documentation}
{manual} {guide} {code examples},
  pdfview={FitV},
  pdffitwindow={true},
  pdfstartview={FitV},
  pdfnewwindow={false},
  pdfdisplaydoctitle={true},
  pdfhighlight={/P},
  plainpages={false},
  unicode={true}
}

% -----------------------------------------------------------------------------
\begin{document}

\title{User Guide for JSBML}
\version{1.0.0 (preview)}

\newcommand{\where}[1]{\,\textsuperscript{#1}}
\newcommand{\divider}[1]{\multicolumn{3}{c}{\emph{#1}:}}

\author{%
  \setlength{\tabcolsep}{20pt}%
  \begin{tabular}{@{}c>{\hspace*{18pt}}cc@{}}
    \divider{Authors}\\[0.75em]
    Andreas Dr\"ager\where{a} & Nicolas Rodriguez\where{b} & Marine Dumousseau\where{b} \\
    Alexander D\"orr\where{a} & Clemens Wrzodek\where{a}   & Finja B\"uchel\where{a} \\
    Florian Mittag\where{a}   & Michael Hucka\where{c}     & \\[2em]
    \divider{Principal investigators}\\[0.75em]
    Nicolas Le Nov\`ere\where{b} & Andreas Zell\where{a}   & Michael Hucka\where{c} \\[2em]
    \divider{Institutional affiliations}\\[0.45em]
  \end{tabular}
  \\
  \where{a\,}Center for Bioinformatics Tuebingen\\
  University of Tuebingen\\
  T\"ubingen, Germany
  \\[1.5em]
  \where{b\,}European Bioinformatics Institute\\
  Wellcome Trust Genome Campus\\
  Hinxton, Cambridge, UK
  \\[1.5em]
  \where{c\,}Computing and Mathematical Sciences\\
  California Insitute of Technology\\
  Pasadena, CA, USA\\
}

\frontNotice{SBML (the Systems Biology Markup Language) is an XML-based
  format for storing and exchanging computational descriptions of
  biological processes.  To read, write, manipulate, and perform
  higher-level operations on SBML files and data streams, software
  applications need to map SBML entities to suitable software objects.
  JSBML provides a pure Java library for this purpose.  It supports all
  Levels and Versions of SBML, and provides many powerful features,
  including facilities to help migrate from the use of libSBML (a popular
  library for SBML that is not written in Java).
  \\ \\
  This document provides an introduction to JSBML and its use.  It is aimed
  at both developers writing new Java-based applications as well as those
  who want to adapt libSBML-based applications to using JSBML.  This user
  guide is a companion to the JSBML API documentation.
  \\ \\
  \centerline{The JSBML home page is \url{http://sbml.org/Software/JSBML/}.}
}


\maketitlepage
\maketableofcontents
\clearpage


\chapter{Getting started with JSBML}
\label{chp:getting-started}

JSBML is a Java\TTra library that will help you to read, write and
manipulate SBML files~\cite{Draeger2011a, Draeger2011b}. This chapter
provides information for quickly getting started with using JSBML.

Before you can use JSBML, you will need to obtain a copy of the library.
\Vref{sec:obtaining-jsbml} below describes different ways of doing
this, and explains which additional libraries you may need. JSBML also
requires the use of a Java Runtime Environment (JRE) version~1.5 or
later~\cite{JavaDownloadURL}. \index{Java Runtime Environment (JRE)} In the
rest of this document, we assume that you have already installed a suitable
JRE or Java Development Kit (JDK), and know how to configure the Java class
path on your system. \index{Java Development Kit (JDK)}

It is also essential to \emph{understand SBML} in order to be able to use
it (and JSBML) properly. If you are not already familiar with SBML, a good
starting point for learning about it is the latest SBML
specification~\cite{SBMLspecs}. You can find answers to many questions in
the SBML FAQ~\cite{SBMLFAQ} and optionally by asking on one of the SBML
discussion lists~\cite{SBMLforums}.

\input{JSBML_getting_started}
\input{JSBML_hello_world_examples}



\chapter{Differences between JSBML and libSBML}
\label{chp:jsbml-libsbml-diffs}

Prior to the availability of JSBML, the most widely-used API library for
SBML offering a Java interface has been libSBML~\cite{Bornstein2008}. As a
result, many Java application developers working with SBML are already
accustomed to the classes, methods and general approach provided by
libSBML. This chapter discusses the main differences between these two
libraries, and is aimed at current libSBML users who want to transition to
using JSBML. We also provide some programming examples and hints about how
to use and work with JSBML.

\input{JSBML_compared_to_libSBML}
\input{JSBML_howto_write_extensions}


\chapter{Acknowledgments}
\label{chp:acknowledgements}

\input{JSBML_Acknowledgments}


\appendix

\chapter{Frequently Asked Questions (FAQ)}
\label{chp:faq}

\input{JSBML_FAQ}


\chapter{Open tasks in JSBML development}
\label{chp:open-tasks}

\input{JSBML_open_tasks}


\clearpage

\thispagestyle{plain}
\pagestyle{plain}
\bibliography{../common/tex/literature}

% Index
\setindexprenote{\vspace*{0.1ex}}
\printindex

\end{document}



% -----------------------------------------------------------------------------
% 2012-04-06 <mhucka@caltech.edu> I think these comments are no longer
% relevant, and can be deleted.  They were originally below the abstract above.


% It includes instructions and descriptions of where and how to obtain the JSBML
% library and the JSBML modules.

%Although the libraries JSBML and libSBML, used to work with files and data
%structures defined in SBML (Systems Biology Markup Language), are
%very similar and share a common scope, users should be informed about their
%major differences to help switch more easily from one library to the other. To
%this end, the document at hand gives a brief overview of the main differences
%between the Java\texttrademark{} application programming interfaces (API) of
%both libraries.
%
%In addition, JSBML can be used as a communication layer between the widespread
%application CellDesigner and any application that works with JSBML as its
%internal data structure. An example is given, that demonstrates how to
%convert between CellDesigner's plugin data structures and JSBML objects.
%
%In the same way, it is possible to inter-convert between data structures
%obtained from libSBML and JSBML \textcolor{red}{No need to mention the
%specific versions in the abstract, I think}. We provides an example of how to
%read SBML files with libSBML, turn them into JSBML data structures, manipulate
% them and turn them back to libSBML for writing.
%
%Furthermore, JSBML will provides a compatibility module, whose member classes
%show an identical API as defined in libSBML. In this way, the
%compatibility module will facilitate a switch from libSBML to JSBML and vice
%versa by simply exchanging the included JAR file in the project.
%
%\textcolor{red}{This document gives an example for the usage of the
%compatibility module.}
% Not sure we need to mention in the abstract that we gives examples code for
% each point and anyway, there is not example at the moment for this module.

% I removed the line return for the \index to try to remove the extra spaces
% that were added.
