For questions regarding SBML, please see the SBML FAQ at
\url{http://sbml.org/Documents/FAQ}.
\begin{description}
\item[\parbox{\textwidth}{Why does the class \texttt{LocalParameter} not inherit
from \texttt{Parameter}?}]
\index{parameter!\texttt{LocalParameter}}
\index{parameter!\texttt{Parameter}}
The reason is the Boolean
\index{Boolean}
attribute \texttt{constant}, which is present in
\index{constant}
\index{parameter!\texttt{constant}}
\texttt{Parameter} and can be set to \texttt{false}. A parameter in the meaning
of SBML is not a constant, it might be some system variable
\index{JSBML!variable@\texttt{Variable}}
and can therefore be the subject of \texttt{Rule}s,
\index{rule}
\texttt{Event}s\index{event!\texttt{Event}}, \texttt{InitialAssignment}s
\index{InitialAssignment@\texttt{InitialAssignment}}
and so on, i.e., all instances of \texttt{Assignment},
\index{JSBML!assignment@\texttt{Assignment}}
whereas a \texttt{LocalParameter} is defined as a constant quantity that never
changes its value during the evaluation of a model\index{model}. It would
therefore only be possible to let \texttt{Parameter} inherit from
\texttt{LocalParameter} but this could lead to a semantic misinterpretation.

\item[\parbox{\textwidth}{Does JSBML depend on SWING or any particular graphical
user interface implementation?}]
Although all classes in JSBML implement the \texttt{TreeNode} interface, which
is located in the package \texttt{javax.swing.tree}, all classes in JSBML are
entirely independent from any graphical user interface, such as the
SWING\index{graphical user interface!swing@\texttt{swing}} implementation. When
loading the \texttt{TreeNode} interface, no other class from SWING will be
initialized or loaded; hence JSBML can also be used on computers that do not
provide any graphical system without the necessity of catching a
\texttt{HeadlessException}. The \texttt{TreeNode} interface only defines
methods and properties that all recursive tree data structures have to
implement anyway. Letting JSBML classes extend this interface makes JSBML
compatible with many other Java classes and methods that make use of the
standard \texttt{TreeNode} interface, hence ensuring a high compatibility with
other Java libraries. Since the SWING package belongs to the standard
Java\texttrademark{} distribution, the \texttt{TreeNode} interface should
always be localized by the Java Virtual Machine, independent from the specific
hardware or system. An exception might Android\index{Android} systems, which
do not provide any parts from the SWING package of Java. Therefore, the JSBML
team is currently developing a specialized \texttt{android} compatibility
module for JSBML.

\item[\parbox{\textwidth}{Does the usage of the the \texttt{java.beans} package
for the \texttt{TreeNodeChangeListener} lead to an incompatibility with light-weight
Java installations?}]
With the \texttt{java.beans} package being part of the standard Java
distribution, such an incompatibility will not occur. Extending existing
standard Java classes leads to a higher compatibility with other libraries and
should therefore be the preferred way to go in the development of JSBML.

\item[\parbox{\textwidth}{Does JSBML support SBML extension packages?}]
In version 0.8, JSBML does not provide an abstract programming interface for
extension packages.%
\index{extension packages}
However, the JSBML community is actively developing extension packages for the
following SBML extensions: \texttt{fba}, \texttt{groups}, \texttt{layout},
\texttt{multi}, \texttt{qual}, and \texttt{spatial}. The release 1.0 of JSBML
will support at least these packages. If you want to make use of SBML
extensions in JSBML before an official release of version 1.0, please checkout
the latest copy from the \texttt{trunk} of 
\ifthenelse{\boolean{includeCodeExample}}{%
  JSBML. To this end, please follow the instructions at
  \url{http://sourceforge.net/projects/jsbml/develop}.
}{
  JSBML (see the description~\vpageref{SVN} for details).
}%

\end{description}
