\documentclass[
  BCOR12mm,
  letterpaper,
  11pt,
  headsepline,
  pointlessnumbers,
  tablecaptionabove,
  onelinecaption,
  headinclude,
  appendixprefix,
  idxtotoc,
  bibtotoc,
  twoside,
  titlepage
]{scrreprt}

\usepackage{JSBMLstyle}

\hypersetup{
  bookmarks={true},
  bookmarksopen={true},
  bookmarksopenlevel={1},
  bookmarksnumbered={true},
  breaklinks={true},
  colorlinks={false},
  pdfpagemode={UseOutlines},
  pdftitle={User guide for JSBML},
  pdfauthor={Andreas Dr\"ager} {Nicolas Rodriguez} { Marine Dumousseau} { Alexander D\"orr} {Clemens Wrzodek} {Nicolas Le Nov{\`e}re} {Andreas Zell} {Michael Hucka},,
  pdfsubject={Software guide},
  pdfkeywords={JSBML} {libSBML} {Java} {SBML} {API} {LaTeX} {documentation} {manual} {guide} {code examples},
  pdfview={FitV},
  pdftex,
  pdffitwindow={true},
  pdfstartview={FitV},
  pdfnewwindow={false},
  pdfdisplaydoctitle={true},
  pdfhighlight={/P},
  plainpages={false},
  unicode={true},
  urlcolor={blue}
}

% for index
\usepackage{makeidx}
\makeindex

\hyphenation{
AST-Node
Boo-lean
Cell-De-sig-n-er
Call-able-S-Base
Event-Listener
Event-Hand-l-er
Event-Object
get-Ex-po-nent
get-Ex-po-nent-As-Double
get-Spatial-Di-men-sions
get-Spatial-Di-men-sions-As-Double
J-SB-ML
Ki-ne-tic-Law
lib-SBML-Constants
Lib-SB-ML-Reader
Local-Pa-ra-meter
Plug-in-Action
plug-in
Property-Change-Event
SBML-Document
Tree-Node
Tree-Node-Change-Event
Tree-Node-Change-Listen-er
Tree-Node-With-Change-Support
time-Units
Tree-Node
}

\title{User guide for JSBML}
\author{Andreas Dr\"ager\thanks{Center for Bioinformatics Tuebingen, University
of Tuebingen, T\"ubingen, Germany}\and%
Nicolas Rodriguez\thanks{European Bioinformatics Institute, Wellcome Trust
Genome Campus, Hinxton, Cambridge, UK}\and%
Marine Dumousseau\footnotemark[2]\and%
Alexander D\"orr\footnotemark[1]\and%
Clemens Wrzodek\footnotemark[1]}
\date{Principal Investigators:\\
Nicolas Le Nov{\`e}re\footnotemark[2],
Andreas Zell\footnotemark[1], and
Michael Hucka\thanks{Computing and Mathematical Sciences, California Institute
of Technology, Pasadena, California, USA}\\[4ex]
\today}

\newboolean{includeCodeExample}
\setboolean{includeCodeExample}{false}

\begin{document}

\maketitle

\begin{abstract}
The specifications of the Systems Biology Markup Language (SBML) define a
standard for storing and exchanging biochemical models in XML-formatted text
files. To perform higher-level operations on these models, e.g., numerical
simulation or visual representation, an appropriate mapping to in-memory objects
is required. To this end, the JSBML library has been developed. JSBML supports
all SBML levels and versions that are available today. In addition, JSBML
provides modules that facilitate the development of CellDesigner plugins or ease
the migration from a libSBML backend.

This document should help you getting started with JSBML. It is not only intended
for users, developing their applications from scratch, but also for users,
switching from libSBML to JSBML.

% It includes instructions and descriptions of where and how to obtain the JSBML
% library and the JSBML modules.


%
%
%Although the libraries JSBML and libSBML, used to work with files and data
%structures defined in SBML (Systems Biology Markup Language), are
%very similar and share a common scope, users should be informed about their
%major differences to help switch more easily from one library to the other. To
%this end, the document at hand gives a brief overview of the main differences
%between the Java\texttrademark{} application programming interfaces (API) of
%both libraries.
%
%In addition, JSBML can be used as a communication layer between the widespread
%application CellDesigner and any application that works with JSBML as its
%internal data structure. An example is given, that demonstrates how to
%convert between CellDesigner's plugin data structures and JSBML objects.
%
%In the same way, it is possible to inter-convert between data structures
%obtained from libSBML and JSBML \textcolor{red}{No need to mention the
%specific versions in the abstract, I think}. We provides an example of how to
%read SBML files with libSBML, turn them into JSBML data structures, manipulate
% them and turn them back to libSBML for writing.
%
%Furthermore, JSBML will provides a compatibility module, whose member classes
%show an identical API as defined in libSBML. In this way, the
%compatibility module will facilitate a switch from libSBML to JSBML and vice
%versa by simply exchanging the included JAR file in the project.
%
%\textcolor{red}{This document gives an example for the usage of the
%compatibility module.}
% Not sure we need to mention in the abstract that we gives examples code for
% each point and anyway, there is not example at the moment for this module.

% I removed the line return for the \index to try to remove the extra spaces
% that were added.

\end{abstract}


\tableofcontents

\chapter{Getting started with JSBML}

The following are quick-start instructions for getting started with JSBML. This
document is based on JSBML version 0.8. Before doing any of the steps below,
you will need to obtain a copy of JSBML either via the SourceForge download
page\footnote{\url{https://sourceforge.net/projects/jsbml/files/jsbml/}} or using
Subversion (SVN) as described below.

\input{tex/JSBML_getting_started}

\input{tex/JSBML_hello_world_examples}

\chapter{Main differences between JSBML and libSBML}

Until today, the Java binding of libSBML has been the main library for
developing Java applications that use SBML. Thus, many Java developers are used
to the methods and commands, libSBML provides. The JSBML team made some effort
to allow those developers a fast and easy switch to this new library. For
example, a libSBML compatibility module has been developed, that implements
existing libSBML methods and simply redirects the parameters to the
corresponding JSBML methods. But it is important that developers, coming from
libSBML, know the main differences between the two libraries. The following
sections give a brief description of those main differences.

\input{tex/JSBML_compared_to_libSBML}

\chapter{Open tasks in JSBML}

\input{tex/JSBML_open_tasks}

\appendix

\chapter{Frequently Asked Questions (FAQ)}
\input{tex/JSBML_FAQ}

\chapter{Acknowledgments and funding}
\input{tex/JSBML_Acknowledgments}


% \clearpage
\bibliographystyle{natbib}
\bibliography{literature}

% Index
\printindex

\end{document}
